Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution \thanks{The work reported here was supported by the Cultural and Educational Grant Agency of Slovak Republic (KEGA) under grant No. KG 014STU-4/2024.}

### 2025-06-22 my_cst/article.tex
strange format

% Required packages for IEEEtran
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

% Additional packages from your article (ensuring they do not interfere with IEEE margins)
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepackage{tikz}
\usepackage{listings}
\usepackage{verbatim}
\usepackage[edges]{forest}
\usepackage{adjustbox}
\usepackage{url}

% Do not load geometry or bookmark in IEEEtran
%\usepackage{geometry}
%\usepackage{bookmark}

@@

\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution\\
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{
\IEEEauthorblockN{Richard Čerňanský}
\IEEEauthorblockA{\textit{Department Name}\\
\textit{Organization}\\
City, Country\\
email@example.com}
\and
\IEEEauthorblockN{Ing. Juraj Petrík}
\IEEEauthorblockA{\textit{Department Name}\\
\textit{Organization}\\
City, Country\\
email@example.com}
}

@@

\begin{IEEEkeywords}
Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code analysis.
\end{IEEEkeywords}

@@

Despite the success of deep learning models in natural language 
processing (NLP), their application to source code authorship attribution on different dataset sizes. 
remains relatively unexplored. Unlike natural language, source code 
exhibits a more rigid syntax and structure, making its representation 
a key challenge. Several approaches have been proposed, ranging from 
direct tokenized representations of the raw source code to abstract 
syntax tree (AST) transformations that capture program logic. However, 
it remains unclear which type of representation is most effective for 
distinguishing authorship across different levels of dataset complexity.

This study systematically compares machine learning and deep learning 
models, including Random Forests, BERT trained on tokenized source code, 
BERT trained on AST representations, and an Attention-based Neural Network 
(AttentionNN) utilizing AST-derived node-to-node paths. The evaluation is 
conducted on a dataset of 3,756 C language functions from Google Code Jam 
(GCJ) spanning 2009 to 2018 \cite{gcj}. Most of the works in the field
predict the authors based on files consisting of multiple language constructs, our 
research however focuses on prediction of author given single function.
The models are assessed based on their accuracy, precision, recall, and training efficiency across different 
author set sizes.

Through this comparison, we aim to determine which model best balances 
accuracy and computational efficiency, as well as the effectiveness of 
different function representations for authorship attribution.

@@

In the paper \cite{information}, authors present division of features based on the method they use for extraction:

@@

@@

\quad Based on the full definition \cite{sun_ast} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
(non-leaf) or leafs. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.

@@

\begin{figure}[!t]
    \centering
    % C code listing in a minipage
    \begin{minipage}{\columnwidth}
    \begin{lstlisting}[caption={Simple C function for addition}, label={lst:add}, basicstyle=\footnotesize\ttfamily, frame=single, numbers=left]
    int add(int a, int b) {
        return a + b;
    }
    \end{lstlisting}
    \end{minipage}

    % AST diagram scaled to column width
    \resizebox{\columnwidth}{!}{%
    \begin{forest}
    for tree={
      font=\fontsize{14pt}{16pt}\selectfont,
      l sep=3cm,
    }
    [TranslationUnit
      [\textbf{FunctionDefinition}: {int add(int a, int b) \{}
        [\textbf{BasicTypeSpecifier}: {int}]
        [\textbf{FunctionDeclarator}: {add(}
          [\textbf{IdentifierDeclarator}: {add}]
          [\textbf{ParameterSuffix}: {(int a, int b)}
            [\textbf{ParameterDeclaration}: {int a}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {a}]
            ]
            [\textbf{ParameterDeclaration}: {int b}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {b}]
            ]
          ]
        ]
        [\textbf{CompoundStatement}: {\{ return a + b; \}}
          [\textbf{ReturnStatement}: {return a + b;}
            [\textbf{AddExpression}: {a + b}
              [\textbf{IdentifierName}: {a}]
              [\textbf{IdentifierName}: {b}]
            ]
          ]
        ]
      ]
    ]
    \end{forest}
    }
    \caption{(a) C function code listing and (b) corresponding AST diagram.}
    \label{fig:ast_struc_example}
\end{figure}

@@

For evaluation of complexity of examined functions we have used Halstead complexity 
measures Difficulty, Volume, Effort \cite{halstead}. To aggregate them into single value, 
we used weighted sum of the three metrics weighting them in ratios 4:3:3 respectively. 
We did not incorporate Halstead length because we did separate analysis on length.

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=3.5cm,height=5cm]{figures/high_level_prediciton_task.png}
    \caption{\cite{compiler_based} High-level overview of the function classification prediction task.}
    \label{fig:predictive_task}
\end{figure}

We utilized various 
types of representations of single function's raw source code to provide the models with data.
These include word tokenized TF-IDF vectors for Random Forests, BPE of raw source code for
BERT, linearized AST pre-order traversal for BERT and representation set of randomly selected 600 node-to-node paths for AttentionNN.  
PsycheC project's C language frontend parser \cite{psychec} was used for parsing the code into ASTs. Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalabillity. The different set sizes were
110, 27, 11, 3 with each having increased demands on minimal function count than the one before. The minimal function counts per author for the sizes were 25, 40, 50, 58 respectively.
Table \ref{tab:basic_props} summarizes the basic properties of our datasets.

\begin{table}[!b]
    \centering
    \caption{Basic properties of our different datasets.}
    \label{tab:basic_props} % label placed AFTER caption
    \begin{tabular}{|c|l|l|}
        \hline
        \textbf{Author set size} & \textbf{No. of functions} & \textbf{Min. function count / author} \\ 
        \hline
        110 & 3756 & 25 \\ 
        \hline
        27 & 1357 & 40 \\ 
        \hline
        11 & 638 & 50 \\ 
        \hline
        3 & 212 & 58 \\ 
        \hline
    \end{tabular}
\end{table}

@@

\begin{table*}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Comparison of models across author attribution accuracy and training time for different author set sizes.}
    \label{tab:comparison_models}
    \centering
    \begin{tabular}{lcccc}
    \hline
    \multicolumn{5}{c}{\textbf{Model Performance on Author Set Sizes}} \\
    \hline
    \textbf{Model} & \textbf{110} & \textbf{27} & \textbf{11} & \textbf{3} \\
    \hline
    Random Forests (TF-IDF) & 66.24\% / 3.63 min & 70.60\% / 0.23 min & 80.56\% / 0.08 min & 85.84\% / 0.05 min \\
    BERT (source code)    & 72.47\% / 30.33 min & 82.72\% / 9.49 min & 89.84\% / 5.30 min  & \textbf{90.70\% / 1.82 min} \\
    BERT (AST traversal)  & 24.07\% / 33.53 min & 35.66\% / 10.30 min & 40.62\% / 6.20 min  & 83.72\% / 1.22 min \\
    AttentionNN (AST paths)& 36.93\% / 3.53 min  & 53.42\% / 1.19 min  & 60.82\% / 0.54 min  & 69.30\% / 0.19 min \\
    \hline
    \end{tabular}
\end{table*}

\begin{figure}[!t]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
        \caption{Frequency distribution of author names}
        \label{fig:distr_names}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
        \caption{Frequency distribution of function lengths}
        \label{fig:disrt_lens}
    \end{subfigure}
    \caption{Important frequency distributions of our dataset.}
    \label{fig:eda_distr}
\end{figure}

From the figure \ref{fig:distr_names} we can see that the most of our dataset's 
functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
most of the time we only have from 10 to 50 tokens available to find the correct author.

@@

In this section we will discuss the training strategies and setup for the examined models.
Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
As training of BERT takes a lot more time than the other 2, we just repeated the traing 3 times and 
observed variance in evaluation metrics lower than 1\%.

@@

For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
Firt we used grid search to optimize these hyperparameters:
\begin{itemize}
    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
    \item \textbf{max\_depth}: \{10, 20, 30, None\}
    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
    \item \textbf{bootstrap}: \{True, False\}
\end{itemize}

and then we trained the model using the best hyperparameters found.

\subsection{BERT}

For both BERT training setups we used Microsoft's pretrained graphcodebert-base \cite{codebert} that already understands general programming concepts like 
control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
is 'fair' across the models that utilize ASTs.

@@

@@

To capture all three of the structural, semantic and syntactic features in a number vector that could be somehow fed into a neural network classifier. 
Code2vec \cite{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
of its components.
Path-contexts capture the node-to-node path in concise manner as it consist of triplets of (start node, the path of nodes kinds, terminal node).
Each distinct node is encoded with the data value it carries and each distinct path is encoded as a string of nodes' kinds in between the start and terminal node.
The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveil the author the most.

@@

    \end{axis}
    \end{tikzpicture}
    \caption{3D bar chart comparing Accuracy, Precision, and Recall across models and author set sizes.}
    \label{fig:3d_bars}
\end{figure*}

@@

As we have mentioned, the models were mainly tested for their scalability with having more functions per author while decreasing the distinct author set size.
Table \ref{tab:comparison_models} summarizes how each model performed in terms of accuracy and the time required for training (until the validation loss stopped decreasing) across diverse datasets
and figure \ref{fig:3d_bars} visualizes the achieved performances on Accuracy, Precision and Recall. From table \ref{tab:comparison_models} we can say that BERT 
trained on tokens extracted from the raw source code performed the best at each of the author set sizes, especially with accuracy over 90\% on the smallest set, however with the longest required time to train the model, more than 
9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less. Our RF experiment on dataset with 110 reached results similar to this study \cite{rf1}, 
where they had 525 authors, but with 2-11 files per author and also included information from comments which we filtered out. In \cite{rf2}, authors reached accuracy of 84.5\% on 50 authors with and average of 42.8 files each similar to our setup for 
27 authors. They scored 14\% more than we did, using the the features presented in \cite{caliskan}.

What also interested us is that From \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function 
count per author from 50 to 58 had little impact on the model's ability to distinguish authors. 
On the other hand, from being the worst, BERT fed with AST pre-order traversal scored a 40\% higher under these conditions. Unfortunately, we were unable to find any relevant sources against which to compare our BERT results.
Nevertheless the following experiments conducted on other models, helped us understand advantages and disadvantages of BERT model for authorship attribution.

The confusion matrix shows \ref{fig:conf_matrix_bert}, that the model performed best on 2 most represented authors
(see \ref{fig:distr_names}). Although if we take 8 most represented authors, only 5 of them are also 
the most accurately classified.

@@

@@

This model showed relatively promising results, indicating increasing dataset size and providing more complex data enhances its performance.
Paper \cite{pbnn} presents similar model reaching 0.979\% accuracy with 40 authors but 3,021 files for each in programming language Java.

Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
we assesed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
Number of nodes in the AST and depth of the AST of the function.

From the figure \ref{fig:ann_acc} we can deduce, there is an emerging trend in accuracy from bins of lower value to the bins of higher value.
Eventhough the frequency distribution of both variables indicates more frequency on the lower values we observe a lower
accuracy. This tells that the model is learning more about the author from structural complex functions than simple non-nesting and 
control-flow divergent ones.

\begin{figure}[!t]
    \centering
    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_lens.pdf}
    \caption{Frequency distribution of misclassified function lengths (in tokens) for Random Forests, BERT, and AttentionNN.}
    \label{fig:misclass_lens}
\end{figure}

>>>>

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\usepackage{tikz}
\usepackage{listings}
\usepackage{verbatim}
\usepackage[edges]{forest}
\usepackage{adjustbox}
\usepackage{url}
\usepackage{float}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\title{Comparing Machine Learning and Deep Learning Models with Attention Mechanisms for Multiple Source Code Authorship Attribution
\thanks{The work reported here was supported by the Cultural and Educational Grant Agency of Slovak Republic (KEGA) under grant No. KG 014STU-4/2024.}
}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Bc. Richard Čerňanský}
\IEEEauthorblockA{
\textit{Faculty of Informatics and Information Technologies} \\
\textit{Slovak University of Technology}\\
Bratislava, Slovakia \\
richard.cernansky3@gmail.com}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Ing. Juraj Petrík}
\IEEEauthorblockA{
\textit{Faculty of Informatics and Information Technologies} \\
\textit{Slovak University of Technology}\\
Bratislava, Slovakia \\
juraj.petrik@stuba.sk}
}

@@

\begin{IEEEkeywords}
    Authorship attribution, machine learning, deep learning, BERT, attention mechanisms, source code analysis
    \end{IEEEkeywords}

@@

Despite the success of deep learning models in natural language 
processing (NLP), their application to source code authorship attribution on different dataset sizes 
remains relatively unexplored. Unlike natural language, source code 
exhibits a more rigid syntax and structure, making its representation 
a key challenge. Several approaches have been proposed, ranging from 
direct tokenized representations of the raw source code to abstract 
syntax tree (AST) transformations that capture program logic. However, 
it remains unclear which type of representation is most effective for 
distinguishing authorship across different levels of dataset complexity.

In the first part we present important related work results in the field of source 
code authorship attribution. Then we continue with explaining methods and techniques 
that we used for developing and evaluating comparative framework of different models.
Section 3 describes the models in more detail.
The main part of this study systematically compares machine learning and deep learning 
models, including Random Forests, BERT trained on tokenized source code, 
BERT trained on AST representations, and an Attention-based Neural Network 
(AttentionNN) utilizing AST-derived node-to-node paths. The evaluation is 
conducted on a dataset of 3,756 C language functions from Google Code Jam 
(GCJ) spanning 2009 to 2018 \citet{petrik}. Most of the works in the field
predict the authors based on files consisting of multiple language constructs, our 
research however focuses on prediction of author given single function.
The models are assessed based on their accuracy, precision, recall, and training efficiency across different 
author set sizes. Through this comparison, we aim to determine which model best balances 
accuracy and computational efficiency, as well as the effectiveness of 
different function representations for authorship attribution.

\section{Related Work}

\begin{table*}[!t]
    \centering
    \includegraphics[width=0.7\linewidth]{figures/dataset_table.png}
    \caption{Table summarizing state-of-the-art works in Source code Authorship Attribution (AA) with machine learning approach \citet{datasets}.}
    \label{tab:datasets_table}
\end{table*}

A number of studies have explored source code authorship attribution using varying 
data representations and model architectures. Early work focused on using statistical
models trained on token frequency vectors or stylometric features. More recent works, 
displayed on the table \ref{tab:datasets_table} show that traditional Machine
Learning (ML) models like Random Forest (RF), Support Vector Machine (SVM) or
Neural Network (NN) can provide reliable results on different sized datasets. 
Specifically, Caliskan et al. \citet{caliskan} used TF-IDF with RF on 
Java code and demonstrated strong attribution capabilities using features extracted 
from abstract syntax trees and code formatting. However, there is no common 
agreement among researchers for selecting the best machine learning classifier 
and data features for source code AA.

In recent years, the development of transformer-based models like Bidirectional Encoder Transformer (BERT) has 
significantly impacted Natural 
Language Processing (NLP), particularly in areas such as Natural language 
understanding and generation (NLG). However, their potential in the field 
of authorship attribution, especially for source code, remains relatively 
underexplored. In some works, they were compared to the state-of-the-art 
models, showing promising results \citet{fi}. We are particularly interested 
in pretrained models such as CodeBERT and GraphCodeBERT \citet{codebert}, 
to source code. These models were initially developed for general code 
understanding tasks (e.g., function naming, defect detection) but have shown 
promising results in Authorship Attribution task when fine-tuned on labeled datasets. 
Choi et. al.~\citet{choi} shows that Large Language Models (LLM) can be successfully utilized in 
code authorship attribution tasks. However, their ultimate utility in 
real-world deployment will struggle with costs, sophisticated prompt engineering, 
model selection, hyperparameter
choices, and an understanding of the models’ sensitivity to
context.

Alon et al. introduced code2vec \citet{code2vec}, an attention-based neural 
network that learns embeddings from paths within Abstract Syntax Tree (AST). 
This architecture enables learning from structural representations of code while 
remaining lightweight compared to transformer models. Attention Neural 
Network (ANN), a similar concept, leverages node-to-node ANN paths to 
predict function authorship. 
Bogomolov et. al~\citet{bogomolov} also examined path-based models on JAVA dataset 
consisting of 200 000 methods and reached 
accuracies around 98\% with Path-based Neural Network and Path-based Random Forests.

However, much of the existing work focuses on large datasets or file-level 
attribution. For example, studies using GitHub datasets typically assume 
access to full repositories or complete files. In contrast, authorship 
attribution at the function level, especially on small datasets such as 
our collection of GCJ solutions, remains unexplored. Moreover, another 
issue lies in the absence of standardized benchmark datasets for the task, 
as the current research relies on datasets with different properties in terms 
of size, author set size, source code background and other.

Moreover, little attention has been paid to the impact of function complexity 
on model performance. Most prior evaluations rely on global accuracy metrics 
without analyzing how model effectiveness varies across functions of differing 
structural or semantic richness.

@@

In the paper \citet{information}, authors present division of features based on the method they use for extraction:

@@

\begin{figure}[!b]

    \centering
    % C code listing in a minipage
    \begin{minipage}{\columnwidth}
        \begin{lstlisting}[
            caption={Simple C function for addition},
            label={lst:add},
            basicstyle=\footnotesize\ttfamily,
            frame=single,
            numbers=left,
            numberstyle=\tiny\color{gray},
            numbersep=5pt,         % distance from code to line number
            xleftmargin=2em        % indent the whole listing 2em from the left
          ]
          int add(int a, int b) {
              return a + b;
          }
          \end{lstlisting}
    \end{minipage}

    % AST diagram scaled to column width
    \resizebox{\columnwidth}{!}{%
    \begin{forest}
    for tree={
      font=\fontsize{14pt}{16pt}\selectfont,
      l sep=3cm,
    }
    [TranslationUnit
      [\textbf{FunctionDefinition}: {int add(int a, int b) \{}
        [\textbf{BasicTypeSpecifier}: {int}]
        [\textbf{FunctionDeclarator}: {add(}
          [\textbf{IdentifierDeclarator}: {add}]
          [\textbf{ParameterSuffix}: {(int a, int b)}
            [\textbf{ParameterDeclaration}: {int a}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {a}]
            ]
            [\textbf{ParameterDeclaration}: {int b}
              [\textbf{BasicTypeSpecifier}: {int}]
              [\textbf{IdentifierDeclarator}: {b}]
            ]
          ]
        ]
        [\textbf{CompoundStatement}: {\{ return a + b; \}}
          [\textbf{ReturnStatement}: {return a + b;}
            [\textbf{AddExpression}: {a + b}
              [\textbf{IdentifierName}: {a}]
              [\textbf{IdentifierName}: {b}]
            ]
          ]
        ]
      ]
    ]
    \end{forest}
    }
        % Figure 3: AST example
    \caption{(a) C function for addition; (b) its Abstract Syntax Tree.}
    \Description{Left panel: numbered C code listing “int add(int a, int b) { return a + b; }”. Right panel: AST tree with nodes for TranslationUnit, FunctionDefinition, BasicTypeSpecifier “int”, FunctionDeclarator “add(…)”, CompoundStatement, ReturnStatement, AddExpression, and IdentifierName leaves “a” and “b”.}
    \label{fig:ast_struc_example}
\end{figure}

@@

\quad Based on the full definition \citet{sun_ast} Abstract syntax tree (AST) is a tree-based representation of source code, where nodes represent various elements of the source code 
and paths represent relationships of the structure. AST is an inseparable part of compilation process for many reasons. It is used as an 
intermediate representation of program utilized for optimization and generation of machine code. Nodes of the tree can be either internal 
non-leaf or leaf nodes. Internal nodes define the program's constructs or operations for their children. Leaf nodes store the actual textual value.

@@

@@

From Figure~\ref{fig:distr_lens}, we observe that functions 
vary both in their length and the same applies also to the structural complexity of their ASTs. 
Therefore, we decided to evaluate the models' performance in terms of function 
complexity. There are various measures of complexity and for our purposes, 
we selected two: the simple length in tokens and a measure called Halstead 
complexity, which quantifies Halstead Difficulty, Volume, and Effort \citet{halstead}. 
To aggregate them into single value, we used weighted sum of the three metrics 
weighting them in ratios 4:3:3 respectively. We chose these ratios because it reflects both the theoretical importance 
and the empirical scale of the three metrics. Halstead Difficulty is most directly 
tied to cognitive complexity, so we give it a slightly larger share (40 \%), 
while Volume (30 \%) and Effort (30 \%) each make substantial but secondary 
contributions. Before applying these weights we normalize each metric 
to the same range, so that the weight split yields a balanced 
composite score—capturing “how hard” (Difficulty), “how big” (Volume), 
and “how much work” (Effort) in one single complexity value. We did not 
incorporate Halstead length because we did separate analysis on length.

@@

We utilized various 
types of representations of single function's raw source code to provide the models with data.
These include word tokenized TF-IDF vectors for Random Forests, BPE of raw source code for
BERT, linearized AST pre-order traversal for BERT and representation set of randomly selected 600 node-to-node paths for AttentionNN.  
PsycheC project's C language frontend parser \citet{psychec} was used for parsing the code into ASTs. Each model was trained in Tensorflow.Keras framework and with different author set sizes to evaluate scalability. The different set sizes were
110, 27, 11, 3 with each having increased demands on minimal function count than the one before. The minimal function counts per author for the sizes were 25, 40, 50, 58 respectively.
Table \ref{tab:basic_props} summarizes the basic properties of our datasets.

\begin{table}[!b]
    \centering
    \caption{Basic properties of our various datasets.}
    \label{tab:basic_props} % label placed AFTER caption
    \begin{tabular}{|c|l|l|}
        \hline
        \textbf{Author set size} & \textbf{No. of functions} & \textbf{Min. function count / author} \\ 
        \hline
        110 & 3756 & 25 \\ 
        \hline
        27 & 1357 & 40 \\ 
        \hline
        11 & 638 & 50 \\ 
        \hline
        3 & 212 & 58 \\ 
        \hline
    \end{tabular}
\end{table}

\begin{figure}[!b]
    \centering
    \includegraphics[width=3.5cm,height=5cm]{figures/high_level_prediciton_task.png}
    \caption{\citet{compiler_based} Pipeline overview for source-code authorship classification.}
    \Description{Flowchart showing raw functions parsed into tokens and ASTs, features extracted (TF-IDF, AST paths, BERT embeddings), and three classifiers (Random Forest, BERT, AttentionNN) producing author predictions.}

    \label{fig:predictive_task}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \begin{subfigure}{0.48\textwidth}
      \includegraphics[width=\linewidth]{figures/distribution_names.pdf}
      \caption{Functions per author.}
      \label{fig:distr_names}
    \end{subfigure}%
    \hfill
    \begin{subfigure}{0.48\textwidth}
      \includegraphics[width=\linewidth]{figures/distr_lens_tokens.pdf}
      \caption{Function length in tokens.}
      \label{fig:distr_lens}
    \end{subfigure}
    \caption{Frequency distributions of functions per author and of function lengths.}
    \Description{Two histograms: left shows the long‐tail distribution of functions across authors; right shows most functions fall between 10 and 50 tokens.}
    \label{fig:eda_distr}
  \end{figure}

From the figure \ref{fig:distr_lens} we can see that the most of our dataset's 
functions lengths lie between 10 and 50 tokens which makes our goal of authorship attribution of single function
harder than if given a whole file consisting of multiple functions or code blocks in sense that for the prediction
most of the time we only have from 10 to 50 tokens available to find the correct author.
After having all of the functions' source codes ready in the suitable form for the model to train we 
performed the training and evaluation on 4 different model setups that are presented in next section.
The high-level overview of the pipeline is illustrated in the figure~\ref{fig:predictive_task}.

@@

In this section we will discuss the training strategies and setup for the examined models.
Random Forests and AttentionNN models were trained using Stratified Cross validation strategy that ensures, that each author is evenly distributed
across the folds. Number of folds was set to 5 and 3 folds were used for training, 1 for validation and 1 for testing.
As training of BERT takes a lot more time than the other 2, we just repeated the training 3 times and 
observed variance in evaluation metrics lower than 1\%.

@@

For comparison purposes, we have chosen Random Forests as standard approach to text based authorship attribution. 
They were trained on word-level tokens using the regex pattern \verb|r"\b\w+\b"|, which matches words separated by word boundary characters.
First we used Grid search to optimize these hyperparameters:
\begin{itemize}
    \item \textbf{n\_estimators}: \{50, 100, 200, 500\}
    \item \textbf{max\_depth}: \{10, 20, 30, None\}
    \item \textbf{min\_samples\_split}: \{2, 5, 10\}
    \item \textbf{min\_samples\_leaf}: \{1, 2, 4\}
    \item \textbf{max\_features}: \{"sqrt", "log2", None\}
    \item \textbf{bootstrap}: \{True, False\}
\end{itemize}

and then we trained the model using the best hyperparameters found. Since each 
dataset was tuned separately via grid search, the optimal hyperparameters vary 
and are not listed here.

\subsection{Bidirectional Encoder Representations from Transformers - BERT}

Transformer-based models such as BERT leverage the self-attention 
mechanism to compute contextual embeddings for each code token by weighting 
interactions across the entire sequence, thus capturing both local syntactic 
patterns and long-range semantic dependencies. Pre-trained on massive corpora 
via masked language modeling and next-sentence prediction, BERT encodes code 
snippets into rich representations summarized by the special \texttt{[CLS]} token. 
By fine-tuning this encoder with a lightweight classification head on labeled 
code–author pairs, the model learns author-specific stylistic cues like naming conventions, formatting habits, 
and preferred idioms.

For both BERT training setups we used Microsoft's pre-trained graphcodebert-base \citet{codebert} that already understands general programming concepts like 
control structures, loops or functions and is able to capture differences between authors faster. We also used the graphcodebert tokenizer which uses 
Byte pair encoding that merges frequently occurring character sequences into tokens, which is appropriate for data like source code of programming language
that has very rigid syntax. The comments in all code samples were filtered out so the attribution task
is 'fair' across the models that utilize ASTs.

@@

\begin{table*}[!ht]
    \renewcommand{\arraystretch}{1.3}
    \caption{Comparison of models across author attribution accuracy and training time for different author set sizes.}
    \label{tab:comparison_models}
    \centering
    \begin{tabular}{lcccc}
    \hline
    \multicolumn{5}{c}{\textbf{Model Performance on Author Set Sizes}} \\
    \hline
    \textbf{Model} & \textbf{110} & \textbf{27} & \textbf{11} & \textbf{3} \\
    \hline
    Random Forests (TF-IDF) & 66.24\% / 3.63 min & 70.60\% / 0.23 min & 80.56\% / 0.08 min & 85.84\% / 0.05 min \\
    BERT (source code)    & 72.47\% / 30.33 min & 82.72\% / 9.49 min & 89.84\% / 5.30 min  & \textbf{90.70\% / 1.82 min} \\
    BERT (AST traversal)  & 24.07\% / 33.53 min & 35.66\% / 10.30 min & 40.62\% / 6.20 min  & 83.72\% / 1.22 min \\
    AttentionNN (AST paths)& 36.93\% / 3.53 min  & 53.42\% / 1.19 min  & 60.82\% / 0.54 min  & 69.30\% / 0.19 min \\
    \hline
    \end{tabular}
\end{table*}

@@

To capture all three of the structural, semantic and syntactic features in a number vector that could be somehow fed into a neural network classifier. 
Code2vec \citet{code2vec} proposes solution design that extracts so called 'path-contexts' which are encoded into vector space using the embedding values 
of its components.
Path-contexts capture the node-to-node path in concise manner as it consist of triplets of (start node, the path of nodes kinds, terminal node).
Each distinct node is encoded with the data value it carries and each distinct path is encoded as a string of nodes' kinds in between the start and terminal node.
The model then learns to distinguish the authors utilizing the attention mechanism, giving larger weights to contexts that help to reveal the author the most.

@@

    \end{axis}
    \end{tikzpicture}
    \caption{Accuracy, precision, and recall across models and author set sizes.}
\Description{3D bar plot where the x-axis lists models (RF, BERT raw, BERT AST, AttentionNN), the y-axis shows author set sizes (110, 27, 11, 3), and the z-axis gives metric values; bars are blue for accuracy, red for precision, and brown for recall.}
    \label{fig:3d_bars}
\end{figure*}

@@

As we have mentioned, the models were mainly tested for their scalability with having more functions per author while decreasing the distinct author set size.
Table \ref{tab:comparison_models} summarizes how each model performed in terms of accuracy and the time required for training (until the validation loss stopped decreasing) across diverse datasets
and figure \ref{fig:3d_bars} visualizes the achieved performances on Accuracy, Precision and Recall. From table \ref{tab:comparison_models} we can say that BERT 
trained on tokens extracted from the raw source code performed the best at each of the author set sizes, especially with accuracy over 90\% on the smallest set, however with the longest required time to train the model, more than 
9.5 times longer than it took to train AttentionNN to its full potential and 36.4 times longer than the time required for Random Forests which scored just 4.8 percent less. Our RF experiment on dataset with 110 reached results similar to this study \citet{rf1}, 
where they had 525 authors, but with 2-11 files per author and also included information from comments which we filtered out. In \citet{rf2}, authors reached accuracy of 84.5\% on 50 authors with and average of 42.8 files each similar to our setup for 
27 authors. They scored 14\% more than we did, using the the features presented in \citet{caliskan}.

What also interested us is that from \texttt{set\_size=11} to \texttt{set\_size=3}, BERT improved only by 0.86\% which is the smallest improvement across all the experiments and indicates that increasing the minimum function 
count per author from 50 to 58 had little impact on the model's ability to distinguish authors. 
On the other hand, from being the worst, BERT fed with AST pre-order traversal scored a 40\% higher under these conditions. Unfortunately, we were unable to find any relevant sources against which to compare our BERT results.
Nevertheless the following experiments conducted on other models, helped us understand advantages and disadvantages of BERT model for authorship attribution.

The confusion matrix shows \ref{fig:conf_matrix_bert}, that the model performed best on 2 most represented authors
(see \ref{fig:distr_names}). Although if we take 8 most represented authors, only 5 of them are also 
the most accurately classified. Additionally, BERT and other models were not biased towards the most represented authors when tested on the 
less represented authors. The only type of bias we have seen was on author that was not in the most represented group. Therefore, we conclude 
that the bias arises not from class imbalance but from the characteristics of each author’s 
functions (e.g., their purpose and functionality).

@@

\begin{figure}[H]
    \centering
    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_lens.pdf}
    % Figure 8: Misclassifications by length
\caption{Misclassification counts by function length for each model.}
\Description{Overlaid line plots (blue=RF, red=BERT, brown=AttentionNN) showing number of misclassified functions binned by token length.}
    \label{fig:misclass_lens}
\end{figure}

\begin{figure}[H]
    \centering
    \hspace*{-25pt}\includegraphics[width=1.2\columnwidth]{figures/misclass_hs.png}
    % Figure 9: Misclassifications by complexity
\caption{Misclassification counts by Halstead complexity for RF and BERT.}
\Description{Overlaid histograms (blue=RF, red=BERT) of misclassified functions versus composite Halstead complexity scores.}
    \label{fig:misclass_halstead}
\end{figure}

@@

This model showed relatively promising results, indicating increasing dataset size and providing more complex data enhances its performance.
Paper \citet{pbnn} presents similar model reaching 0.979\% accuracy with 40 authors but 3,021 files for each in programming language Java.

Considering the observation in figure \ref{fig:misclass_lens}, where AttentionNN performed worst on lengths between 20 and 40 tokens,
we assessed the model in a deeper manner. Evaluating accuracies across functions distributed into bins according to 2 metrics. 
Number of nodes in the AST and depth of the AST of the function.
